Probablistic Model for High Precision/Recall Feature Selection

### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/derivations/Derivation.tex

\subsection{Recall case}
In order to select the subset of features providing high recall binary classifier we need at least one $y^d_i$ to predict true. Therefore, we need a disjunction operation between the predictors $y^d_i$. That is, when $y^d$ is true, we need at least one predictor $y^d_i$ to equals to true, and if $y^d$ were false, all of the predictors $y^d_i$ would have to be false. Then the probability $P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)$ can be expressed as follows:
\begin{equation}
P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)=I\left[y^d=\bigvee_{i=1}^k y^d_i\right]=
\begin{cases}
	y^d=1 \text{ when } \{y_i^d\}_{1\leq i\leq k,\exists{y_j^d\|y_j^d=1}}\\
	y^d=0 \text{ when } \{y_i^d=0\}_{1\leq i\leq k}
\end{cases}\label{eq:expect_recall_1}
\end{equation}
Here, we combined equations \eqref{eq:expect} and \eqref{eq:expect_recall_1}, separated each term according to the actual label value $y^d$ and used the probability sum rule to rewrite the second term as follows:
\begin{align}
f_k^*&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}I\left[y^d=\bigvee_{i=1}^k y^d_i\right]\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)\nonumber\\[0.5em]
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}
\begin{cases}
	y^d=0: \prod_{i=1}^kP\left(y_i^d=0|\vec{x}^{\,d},f_i\right)\\
	y^d=1: \sum_{\substack{\{y_i^d\}_{1\leq i\leq k},\\ \exists{y_i^d\|y_i^d=1}}}\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)
\end{cases}\nonumber\\[0.5em]
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}
\begin{cases}
	y^d=0: \prod_{i=1}^kP\left(y_i^d=0|\vec{x}^{\,d},f_i\right)\\
	y^d=1: 1-\prod_{i=1}^kP\left(y_i^d=0|\vec{x}^{\,d},f_i\right)
\end{cases}\label{eq:expect_recall_2}\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}
\begin{cases}
	y^d=0: \prod_{i=1}^kP\left(y^d=0|\vec{x}_i^{\,d},f_i\right)\\
	y^d=1: 1-\prod_{i=1}^kP\left(y^d=0|\vec{x}_i^{\,d},f_i\right)
\end{cases}\nonumber
\end{align}
From \eqref{eq:expect_recall_2} we can intuitively describe how this equation is related to the precision metric. In the confusion matrix, Recall is calculated dividing the numbers of actual true positives by the number of all true positives (true positives and false negatives). When $y^d$ is false, \eqref{eq:expect_recall_2} gives higher score to feature that have higher probability to predict false, encouraging true negatives. Meanwhile, if $y^d$ is true, \eqref{eq:expect_recall_2} gives lower score to features that have higher probability to predict false, penalizing false negatives.

@@

\section{Likelihood}
\begin{align}
f_k^*&\equiv\operatorname*{arg\,max}_{f_k}\,log\prod_{d\in D}P\left(y^d|\vec{x}^{\,d},F_k\right)\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}log\left(\frac{P\left(y^d,\vec{x}^{\,d},F_k\right)}{P\left(\vec{x}^{\,d},F_k\right)}\right)\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}log\sum_{\{y_i^d\}_{1\leq i\leq k}}\frac{P\left(y^d,\vec{x}^{\,d}, F_k,\{y_i^d\}_{1\leq i\leq k}\right)}{P\left(\vec{x}^{\,d},F_k\right)}\nonumber\\
&=\operatorname*{arg\,max}_{f_k} \sum_{d \in D} log\sum_{{\{y_i^d\}}_{1\leq i \leq k}} \frac{P\left(y^d| \{y^d_i\}_{1 \leq i \leq k}\right) \prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)P\left(\vec{x}^{\,d}\right)\prod_{i=1}^kP\left(f_i\right)}{P\left(\vec{x}^{\,d}, F_k\right)}\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}log\sum_{{\{y_i^d\}}_{1\leq i\leq k}}P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)\nonumber
\end{align}
Here, we rewrote the likelihood of a binary event as its probability, factorized the conditional probability in the joint probability divided by the marginal, marginalized over $\{y^d_i\}_{1\leq i\leq k}$, factorized joint probability in conditional and prior following the graphical model and exploited d-separation to remove irrelevant conditions and to cancel terms in the equation.

Since $\sum_{d\in D}log\sum_{{\{y_i^d\}}_{1\leq i\leq k}}P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)$ is exactly the same as the result in the expectation case, and the logarithm is a monotonous function, the choice of likelihood instead of expectation has no further affect on the algorithm.
\end{document}

>>>>

\subsection{Recall}
In order to select the subset of features providing high recall binary classifier we need at least one $y^d_i$ to predict true. Therefore, we need a disjunction operation between the predictors $y^d_i$. That is, when $y^d$ is true, we need at least one predictor $y^d_i$ to equals to true, and if $y^d$ were false, all of the predictors $y^d_i$ would have to be false. Then the probability $P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)$ can be expressed as follows:
\begin{equation}
P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)=I\left[y^d=\bigvee_{i=1}^k y^d_i\right]=
\begin{cases}
	y^d=1 \text{ when } \{y_i^d\}_{1\leq i\leq k,\exists{y_j^d\|y_j^d=1}}\\
	y^d=0 \text{ when } \{y_i^d=0\}_{1\leq i\leq k}
\end{cases}\label{eq:expect_recall_1}
\end{equation}
Here, we combined equations \eqref{eq:expect} and \eqref{eq:expect_recall_1}, separated each term according to the actual label value $y^d$ and used the probability sum rule to rewrite the second term as follows:
\begin{align}
f_k^*&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}I\left[y^d=\bigvee_{i=1}^k y^d_i\right]\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)\nonumber\\[0.5em]
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}
\begin{cases}
	y^d=0: \prod_{i=1}^kP\left(y_i^d=0|\vec{x}^{\,d},f_i\right)\\
	y^d=1: \sum_{\substack{\{y_i^d\}_{1\leq i\leq k},\\ \exists{y_i^d\|y_i^d=1}}}\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)
\end{cases}\nonumber\\[0.5em]
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}
\begin{cases}
	y^d=0: \prod_{i=1}^kP\left(y_i^d=0|\vec{x}^{\,d},f_i\right)\\
	y^d=1: 1-\prod_{i=1}^kP\left(y_i^d=0|\vec{x}^{\,d},f_i\right)
\end{cases}\label{eq:expect_recall_2}\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}
\begin{cases}
	y^d=0: \prod_{i=1}^kP\left(y^d=0|\vec{x}_i^{\,d},f_i\right)\\
	y^d=1: 1-\prod_{i=1}^kP\left(y^d=0|\vec{x}_i^{\,d},f_i\right)
\end{cases}\nonumber
\end{align}
From \eqref{eq:expect_recall_2} we can intuitively describe how this equation is related to the precision metric. In the confusion matrix, Recall is calculated dividing the numbers of actual true positives by the number of all true positives (true positives and false negatives). When $y^d$ is false, \eqref{eq:expect_recall_2} gives higher score to feature that have higher probability to predict false, encouraging true negatives. Meanwhile, if $y^d$ is true, \eqref{eq:expect_recall_2} gives lower score to features that have higher probability to predict false, penalizing false negatives.

@@

\section{Likelihood}
\begin{align}
f_k^*&\equiv\operatorname*{arg\,max}_{f_k}\prod_{d\in D}P\left(y^d|\vec{x}^{\,d},F_k\right)\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\,log\prod_{d\in D}P\left(y^d|\vec{x}^{\,d},F_k\right)\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}log\left(\frac{P\left(y^d,\vec{x}^{\,d},F_k\right)}{P\left(\vec{x}^{\,d},F_k\right)}\right)\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}log\sum_{\{y_i^d\}_{1\leq i\leq k}}\frac{P\left(y^d,\vec{x}^{\,d}, F_k,\{y_i^d\}_{1\leq i\leq k}\right)}{P\left(\vec{x}^{\,d},F_k\right)}\nonumber\\
&=\operatorname*{arg\,max}_{f_k} \sum_{d \in D} log\sum_{{\{y_i^d\}}_{1\leq i \leq k}} \frac{P\left(y^d| \{y^d_i\}_{1 \leq i \leq k}\right) \prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)P\left(\vec{x}^{\,d}\right)\prod_{i=1}^kP\left(f_i\right)}{P\left(\vec{x}^{\,d}, F_k\right)}\nonumber\\
&=\operatorname*{arg\,max}_{f_k}\sum_{d\in D}log\sum_{{\{y_i^d\}}_{1\leq i\leq k}}P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)\nonumber
\end{align}
Here, we rewrote the likelihood of a binary event as its probability, factorized the conditional probability in the joint probability divided by the marginal, marginalized over $\{y^d_i\}_{1\leq i\leq k}$, factorized joint probability in conditional and prior following the graphical model and exploited d-separation to remove irrelevant conditions and to cancel terms in the equation.

Since $\sum_{d\in D}log\sum_{{\{y_i^d\}}_{1\leq i\leq k}}P\left(y^d|\{y^d_i\}_{1\leq i\leq k}\right)\prod_{i=1}^kP\left(y_i^d|\vec{x}^{\,d},f_i\right)$ is exactly the same as the result in the expectation case except the logarithm, which is a monotonous function, the choice of likelihood instead of expectation has no further affect on the algorithm.
\end{document}
